---
title: "Exercises"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Section 7.2 exercises

1.  What function would you use to read a file where fields were separated with "\|"?\
    A: with `read_delim(file, "|")`.

2.  Apart from `file`, `skip`, and `comment`, what other arguments do `read_csv()` and `read_tsv()` have in common?

3.  What are the most important arguments to `read_fwf()`?\
    A: `col_positions` and `col_types`.

4.  Sometimes strings in a CSV file contain commas.
    To prevent them from causing problems, they need to be surrounded by a quoting character, like `"` or `'`. By default, `read_csv()` assumes that the quoting character will be `"`.
    To read the following text into a data frame, what argument to `read_csv()` do you need to specify?\
    A:
    ```{r}
    read_csv("x,y\n1,'a,b'", quote = "\'")
    ```

5.  Identify what is wrong with each of the following inline CSV files.
    What happens when you run the code?

    ```{r}
    #| eval: false
    read_csv("a,b\n1,2,3\n4,5,6")
    read_csv("a,b,c\n1,2\n1,2,3,4")
    read_csv("a,b\n\"1")
    read_csv("a,b\n1,2\na,b")
    read_csv("a;b\n1;3")
    ```

6.  Practice referring to non-syntactic names in the following data frame by:

    a.  Extracting the variable called `1`.
    b.  Plotting a scatterplot of `1` vs. `2`.
    c.  Creating a new column called `3`, which is `2` divided by `1`.
    d.  Renaming the columns to `one`, `two`, and `three`.

    ```{r}
    annoying <- tibble(
      `1` = 1:10,
      `2` = `1` * 2 + rnorm(length(`1`))
    )
    
    annoying |> 
      select(`1`)
    
    annoying |> 
      ggplot(
        aes(x = `1`, y = `2`)
      ) + 
      geom_point()
    
    annoying |> 
      mutate(
        `3` = `2` / `1`
      ) |> 
      rename(
        one = `1`, 
        two = `2`, 
        three = `3`
      )
    ```
