---
title: "Exercises"
output: html_document
---

```{r setup, include=FALSE}
library(nycflights13)
library(tidyverse)
```

## Section 3.2 exercises

1.  In a single pipeline for each condition, find all flights that meet the condition:
    -   Had an arrival delay of two or more hours
    -   Flew to Houston (`IAH` or `HOU`)
    -   Were operated by United, American, or Delta
    -   Departed in summer (July, August, and September)
    -   Arrived more than two hours late but didn't leave late
    -   Were delayed by at least an hour, but made up over 30 minutes in flight\
A:\
    - 
    ```{r}
    flights |> filter(dep_delay >= 120)
    ```
    - 
    ```{r}
    flights |> filter(dest %in% c("IAH", "HOU"))
    ```
    - 
    ```{r}
    flights |> filter(carrier %in% c("UA", "AA", "DL"))
    ```
    - 
    ```{r}
    flights |> filter(month %in% c(7, 8, 9))
    ```
    - 
    ```{r}
    flights |> filter(dep_delay <= 0 & arr_delay > 120)
    ```
    - 
    ```{r}
    flights |> filter(dep_delay >= 60 & dep_delay - arr_delay >= 30)
    ```

2.  Sort `flights` to find the flights with the longest departure delays.
    Find the flights that left earliest in the morning.\
    A: 
    ```{r}
    flights |> arrange(desc(dep_delay))
    flights |> arrange(dep_time)
    ```

3.  Sort `flights` to find the fastest flights.
    (Hint: Try including a math calculation inside of your function.)\
    A:
    ```{r}
    flights |> arrange(desc(distance / air_time))
    ```

4.  Was there a flight on every day of 2013?\
    A: Yes.
    ```{r}
    flights |> distinct(month, day)
    ```

5.  Which flights traveled the farthest distance?
    Which traveled the least distance?
    ```{r}
    flights |> arrange(desc(distance))
    flights |> arrange(distance)
    ```

6.  Does it matter what order you used `filter()` and `arrange()` if you're using both?
    Why/why not?
    Think about the results and how much work the functions would have to do.\
    A: The resulting data frame should be the same, but if we apply `arrange()` first, it'll need to sort through rows that are only to be discarded later by `filter()`. Applying `filter()` first would be more efficient.

# Section 3.3 exercises

1.  Compare `dep_time`, `sched_dep_time`, and `dep_delay`.
    How would you expect those three numbers to be related?\
    A: `dep_delay == (dep_time %% 100 - sched_dep_time %% 100) + (dep_time %/% 100 - sched_dep_time %/% 100) * 60`

2.  Brainstorm as many ways as possible to select `dep_time`, `dep_delay`, `arr_time`, and `arr_delay` from `flights`.\
    A: 
    ```{r}
    flights |> select(starts_with(c("dep", "arr")))
    ```

3.  What happens if you specify the name of the same variable multiple times in a `select()` call?\
    A: multiple same variables register only once in the output of `select()`.
    ```{r}
    flights |> select("tailnum", "tailnum")
    ```

4.  What does the `any_of()` function do?
    Why might it be helpful in conjunction with this vector?

    ```{r}
    variables <- c("year", "month", "day", "dep_delay", "arr_delay")
    ```
    A: `any_of()` doesn't check for missing values, i.e., it doesn't throw an error when a variable specified isn't present in the data frame.

5.  Does the result of running the following code surprise you?
    How do the select helpers deal with upper and lower case by default?
    How can you change that default?
    ```{r}
    flights |> select(contains("TIME"))
    ```
    A: quite surprising that `contains()` defaults to ignoring upper and lower cases. setting the argument `ignore.case` to `FALSE` helps `contains()` distinguish cases.
    ```{r}
    flights |> select(contains("TIME", ignore.case = FALSE))
    ```

6.  Rename `air_time` to `air_time_min` to indicate units of measurement and move it to the beginning of the data frame.\
    A: 
    ```{r}
    flights |> relocate(air_time_min = air_time)
    ```

7.  Why doesn't the following work, and what does the error mean?

    ```{r}
    #| error: true
    flights |> 
      select(tailnum) |> 
      arrange(arr_delay)
    ```
    A: variable `arr_delay` doesn't exist anymore after selecting only `tailnum`.

## Section 3.5 exercises

1.  Which carrier has the worst average delays?
    Challenge: can you disentangle the effects of bad airports vs. bad carriers?
    Why/why not?
    (Hint: think about `flights |> group_by(carrier, dest) |> summarize(n())`)\
    A: We might not be able to differentiate whether delays were caused by airports or carriers, because carrier and airports are not independent variables (certain carriers may deal with flights to certain destinations more than others.)
    ```{r}
    flights |> 
      summarize(
        avg_delay = mean(dep_delay, na.rm = TRUE), 
        .by = carrier
      ) |>
      arrange(desc(avg_delay))
    ```

2.  Find the flights that are most delayed upon departure to each destination.\
    A: 
    ```{r}
    flights |> 
      group_by(dest) |>
      slice_max(dep_delay, n = 1) |>
      relocate(dest)
    ```

3.  How do delays vary over the course of the day?
    Illustrate your answer with a plot.\
    A: 
    ```{r}
    avg_delays_by_hour <- flights |>
      select(dep_delay, arr_delay, hour) |>
      summarize(
        avg_dep_delay = mean(dep_delay, na.rm = TRUE), 
        avg_arr_delay = mean(arr_delay, na.rm = TRUE), 
        .by = hour
      )
    ggplot(avg_delays_by_hour, aes(x = hour, y = avg_dep_delay)) + 
      geom_line()
    ggplot(avg_delays_by_hour, aes(x = hour, y = avg_arr_delay)) + 
      geom_line()
    ```

4.  What happens if you supply a negative `n` to `slice_min()` and friends?\
    A: a negative value of `n` will be subtracted from the group size.
    ```{r}
    flights |>
      group_by(month) |>
      slice_min(dep_delay, n = 1)
    flights |>
      group_by(month) |>
      slice_min(dep_delay, n = -1)
    ```

5.  Explain what `count()` does in terms of the dplyr verbs you just learned.
    What does the `sort` argument to `count()` do?\
    A: `count()` can be implemented by `group_by()` and `n()`. The `sort` argument if set to `TRUE`, sorts the rows descendingly by count.
    ```{r}
    flights |>
      count(month)
    flights |> 
      group_by(month) |>
      summarize(n = n())
    ```

6.  Suppose we have the following tiny data frame:

    ```{r}
    df <- tibble(
      x = 1:5,
      y = c("a", "b", "a", "a", "b"),
      z = c("K", "K", "L", "L", "K")
    )
    ```

    a.  Write down what you think the output will look like, then check if you were correct, and describe what `group_by()` does.

        ```{r}
        df |>
          group_by(y)
        ```

    b.  Write down what you think the output will look like, then check if you were correct, and describe what `arrange()` does.
        Also, comment on how it's different from the `group_by()` in part (a).

        ```{r}
        df |>
          arrange(y)
        ```

    c.  Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does.

        ```{r}
        df |>
          group_by(y) |>
          summarize(mean_x = mean(x))
        ```

    d.  Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does.
        Then, comment on what the message says.

        ```{r}
        df |>
          group_by(y, z) |>
          summarize(mean_x = mean(x))
        ```

    e.  Write down what you think the output will look like, then check if you were correct, and describe what the pipeline does.
        How is the output different from the one in part (d)?

        ```{r}
        df |>
          group_by(y, z) |>
          summarize(mean_x = mean(x), .groups = "drop")
        ```

    f.  Write down what you think the outputs will look like, then check if you were correct, and describe what each pipeline does.
        How are the outputs of the two pipelines different?

        ```{r}
        df |>
          group_by(y, z) |>
          summarize(mean_x = mean(x))
            
        df |>
          group_by(y, z) |>
          mutate(mean_x = mean(x))
        ```

